Colonies - Compute Continuums across Platforms

### 2023-04-13 paper_2023/paper.tex
second draft

To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms. Colonies enables development of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.

@@

This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can easily integrate with any third-party system. Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment. The primary objective is to establish a platform where workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet. The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across platforms.

@@

The advantage of Colonies for implementing a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system. For example, if an executor is missing, the process is simply enqueued until a free executor becomes available. A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system. In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption, or minimize network latency by utilizing edge servers.

\subsection{Edge computing}

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.43]{edge.png}
	\caption{Using Colonies to control drones.}
	\label{fig:dashboard2}
\end{figure}

@@

Figure \ref{fig:hpc_workflow} shows an example how Colonies could be used in a HPC environment. Several executors run at the HPC system, each with their own responsibilities. The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g. Singularity containers\footnote{https://docs.sylabs.io}. The Data Importer Executors downloads data to the HPC file system. The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.

@@

\section{Discussion}
This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms. Colonies enable the creation of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems. Instead of directly connecting different systems togeather, Colonies allows machines or users to publish instructions and then assign the instructions to other machines. By chaining these instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.

@@

A significant portion of the paper has focused on describing how to handle robustness and allow software components to fail without compromising the entire system. A critical design decision that makes Colonies robust and scalable is its stateless architecture. Colonies also makes sure that workflow execution run to completion, even if some executor failes. This makes it possible to implement robust data pipelines, but also to efficiently handle IT operations such as Continuous Deployments (CD) and software upgrades. Both security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across platforms and handle constantly changing infrastructures.

It has been mentioned in the paper several times that Colonies is designed to simplify integration with other systems, but what feature in Colonies makes integration easier? Microservices is a design principle used by Colonies to simplify integration. As executors are designed to be loosly-coupled and can be implemented in any language (supported by Colonies), the whole system becomes easier to develop and maintain. For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of complex workflows running across platforms. By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors. This also opens up an exciting possibility to delegate colony management responsibility to another software to implement a serverless computing platform capable of automatically provisioning and deploying executors whenever they are needed.

The paper has also presented some potential use cases. In all of the use cases, Colonies serves as intermediary layer that breaks down complex heterogenous systems in to smaller, resilient and more scalable services, and then offer a unified interface for the services to interact with each other. For example, in HPC use case, Colonies can help to simplify HPC workflow managemewnt by breaking down complex perations into a set of functions that can be executed as part of a workflow.

% The paper has presented several potential use cases where Colonies serves as an intermediary layer that decomposes complex heterogeneous systems into smaller, resilient, and scalable services, and then provide a unified interface for seamless interaction between these services. For instance, in the HPC use cases, Colonies can streamline the workflow management process by HPC operations into a collection of functions that can be seamlessly executed as part of a workflow.

>>>>

To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms. Colonies enables development of a \emph{decentralized runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.

@@

The primary objective of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another. To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to integrate with third-party systems to create a seamlessly integrated computing ecosystem. Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment. The primary objective is to establish a platform where workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet. The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across platforms.

@@

The advantage of using Colonies to implement a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system. For example, if an executor is missing, the process is simply enqueued until a free executor becomes available. A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system. In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption, or minimize network latency by utilizing edge servers.

\subsection{Edge computing}
Edge computing is a distributed computing model that brings computing resources closer to the source where data is generated to improve efficiency or reduce latency. In edge computing, computing and processing is usually done on the \emph{edge} of a network, such as at a server connected to a 5G base station. One could for example imagine a drone video surveillance use case where parts of a video processing pipeline run on an edge server. This approach would make it possible to execute complex AI models that would otherwise be difficult to run on a drone with limited computing capabilities.

Setting up a video pipeline is a complex problem. First, the drone must be able to discover nearby edge servers, make sure the edge server has sufficient resources, deploy some software components at the edge server, then start a video server and finally connect to the video server. Orchestrating such a workflow requires configuration and provisioning of computational resources across platforms.

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.43]{edge.png}
	\caption{Edge workflows, using Colonies to control drones.}
	\label{fig:edge}
\end{figure}

The Colonies framework has the potential to streamline the setup of edge computing workflows by enabling devices and edge servers to be represented as distributed executors that can interact with other. Figure \ref{fig:edge} illustrates a setup of three executors. The \emph{Edge Executor} is responsible for deploying and setting up video processing software at an edge server, while the \emph{Drone Executor} is responsible for activating the video camera at the drone and establishing a connection to the video server hosted on the edge server. The \emph{Video Surveillance Executor} is responsible for identifying suitable edge servers and facilitating submission of workflows by users. Discovery of edge executors can be achieved by interacting with a Colonies server and searching the colony for available executors.

Figure \ref{fig:edge} also depicts a Colonies workflow. The workflow guarantees that the video pipeline is set up in the right order. For example, the drone executor should only connect to the edge server after the video server is running. The workflow can also be used to pass information between the executors. After the \emph{Edge Executor} has set up a video server, it sends a \emph{close} request containing the IP address to the video server. The IP address will then be passed as input to the next process in the workflow, allowing the \emph{Drone Executor} to establish a connection to the edge server.

@@

Figure \ref{fig:hpc_workflow} shows an example how Colonies could be used in a HPC environment. Several executors run at the HPC system, each with their own responsibilities. The \emph{Container Importer Executor} is responsible for converting container images to a format that is supported by Slurm, e.g. Singularity containers\footnote{https://docs.sylabs.io}. The \emph{Data Importer Executor} downloads data to the HPC file system. The \emph{Slurm Executor} is responsible for submitting and overseeing Slurm jobs, while the \emph{Data Exporter} executor is responsible for uploading result data to the cloud.

@@

\section{Discussion}
This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms. Colonies enable the creation of a \emph{decentralized runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems. Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services. By chaining instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.

@@

A significant portion of the paper has focused on describing how to handle robustness to allow software components to fail without compromising the entire system. A critical design decision that makes Colonies robust and scalable is its stateless architecture. Colonies also make sure that workflow execution runs to completion, even if some executors fail. This makes it possible to implement robust data pipelines, but also efficiently handle IT operations such as continuous deployments and software updates. Both security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across platforms and handling constantly changing infrastructures.

It has been mentioned in the paper several times that Colonies is designed to simplify integration with other systems, but what feature in Colonies makes integration easier? Microservices is a design principle used by Colonies to simplify integration. As executors are designed to be loosely-coupled microservices, the entire system becomes easier to develop, test, and maintain. For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of complex workflows running across platforms. By designing each executor as a standalone microservice, it also becomes easy to scale the system by simply deploying more executors. This also opens up for an intriguing possibility to delegate colony management responsibility to a software agent that can automatically provision and deploy executors on demand.

While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a decentralized microservice architecture that allows microservices to reside anywhere on the Internet. Also, by acting as an intermediary layer, Colonies provides a unified interface to facilitate interaction between these microservices, making it possible to create workflows that span multiple platforms. For instance, Colonies could be used to expose an entire HPC system as a microservice and then connected it to other microservices running in the cloud or IoT devices.

In summary, the primary objective of Colonies is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another. Achieving seamless compute continuum requires development of simple, secure, scalable, and robust solutions. Colonies serves as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
